---
title: "Two-Arm Group-Sequential-Design with Binary endpoints"
---

```{r}
#| label: setup
source("../quarto-setup.R")
```

## Design space explored

All comparisons were performed for a **two‑arm, group‑sequential randomized clinical trial** with a **binary primary endpoint**. The aim was to evaluate the consistency of **sample size calculations** across commonly used software tools, over the same factorial design space used in the fixed‑design setting.

### Varying design parameters

The same four design parameters as in the fixed-design section were varied:

-   **Type I error rate ($\alpha$, two-sided):**  
    $\alpha \in {0.01,\ 0.05,\ 0.10,\ 0.20,\ 0.49}$

-   **Power ($1 - \beta$):**  
    $1 - \beta \in {0.51,\ 0.80,\ 0.90,\ 0.99}$

-   **Control-group response rate:**  
    $\pi_c \in {0.10,\ 0.30,\ 0.50,\ 0.80,\ 0.90}$

-   **Treatment effect (difference in proportions):**  
    $\Delta\pi \in {0.05,\ 0.15,\ 0.25,\ 0.49}$

The experimental-group rate was defined as  
$\pi_e = \pi_c + \Delta\pi$,  
and scenarios with $\pi_e > 1$ were excluded.

A total of **300 valid design scenarios** were obtained.

### Group-sequential design assumptions

To isolate differences attributable to software implementation and numerical methods, the group-sequential structure was fixed across all scenarios:

-   **Number of analyses:** 4 looks
-   **Spacing:** equally spaced by **information rate**
-   **Type I error spending:**  
    **Lan–DeMets O’Brien–Fleming–like** α‑spending function
-   **Power spending (futility):**  
    **Lan–DeMets O’Brien–Fleming–like** β‑spending, **non‑binding**
-   **Test:** two-sided test for a **difference in proportions (pooled Z-statistic)**
-   **Allocation ratio:** 1 : 1

Stopping for futility followed a **non‑binding** rule so that crossing the futility boundary does not inflate the Type I error.

### Relevancy classification of design scenarios

The same **high / medium / low relevance** classification used in the fixed-design case was applied:

-   **High relevance** scenarios reflect typical late-phase binary-endpoint designs  
    (e.g. $\alpha \le 0.10$, power 80–90%, control rate between 0.2 and 0.6, effect sizes of 10–20 percentage points).

-   **Medium relevance** scenarios represent extreme but plausible configurations,  
    such as very low or very high baseline rates, or unusually large effect sizes.

-   **Low relevance** scenarios include combinations rarely used in practice  
    (e.g. $\alpha = 0.49$, $\Delta\pi = 0.49$ on a high baseline rate, or power close to 50% or 100%).  
    These scenarios were retained intentionally to assess numerical stability and robustness of asymptotic approximations.

This classification was not used to restrict the analysis; it is only meant to contextualize discrepancies and identify when differences reflect methodological subtleties rather than clinically meaningful effects.

## Results

Method compared:

- East: Difference of Proportions test [PN-2S-DI]
- `rpact::getSampleSizeRates(design = rpact::getDesignGroupSequential)`


```{r}
x_pooled <- read_rds("../outputs/comp_bin_gs_pooled.rds")
```

::: {.light-content}
```{r}
#| out-width: 100%
gt_theme_flatly(x_pooled$tables$table_n_ratio)
x_pooled$plots$p_n_ratio + theme_flatly()
```
:::

::: {.dark-content}
```{r}
#| out-width: 100%
gt_theme_darkly(x_pooled$tables$table_n_ratio) 
x_pooled$plots$p_n_ratio + theme_darkly()
```
:::

- Across **all relevancy levels**, `rpact` shows **excellent agreement** with East, with N‑ratios clustering tightly around 1.00.
- In **high** and **medium‑relevancy** scenarios, variation is minimal: nearly all designs fall within a very narrow band (≈0.98–1.02), indicating practical equivalence.
- In the **low‑relevancy** region, a small number of mild deviations appear, but no systematic trend is observed. These differences stem from edge‑case parameter combinations rather than method behaviour.

**Overall:** `rpact` reproduces East’s group‑sequential Z‑pooled binary sample sizes with **near‑perfect consistency**, including in extended design regions.

### Interactive results

You can explore all results interactively, filter designs, compare methods, and inspect individual cases in the interactive section of the website.

**➡️ [*“Interactive results”*](../interactive.qmd)**