---
title: "One-Arm Fixed Design with Survival endpoints"
bibliography: ../../references.bibtex
---
```{r}
#| label: setup
source("../results-setup.R")
```

## Design space explored

All comparisons were conducted for a **one-arm, fixed-design randomized clinical trial** with a **survival primary endpoint**, analyzed using the logrank test under a proportional hazards framework. The objective was to compare sample size and event calculations across commonly used software implementations over a broad but structured design space.

### Varying design parameters

Four key design parameters were varied systematically:

- **Type I error rate ($\alpha$, two-sided):**  
  $\alpha \in \{0.01, 0.05, 0.10, 0.20, 0.49\}$

- **Power ($1 - \beta$):**  
  $1 - \beta \in \{0.51, 0.80, 0.90, 0.99\}$

- **Hazard ratio (HR):**  
  $\text{HR} \in \{0.10, 0.50, 0.70, 0.90, 0.99\}$

- **Survival probability at 3 years under null hypothesis:**  
  $S_0(3\text{y}) \in \{0.10, 0.30, 0.60, 0.90\}$

The 3-year survival probability was used to calibrate the baseline hazard under an exponential survival assumption, ensuring consistency across methods requiring explicit specification of the underlying survival distribution.

The full factorial combination of these parameters yielded **400 distinct design scenarios**.

### Fixed design assumptions

```{r}
p_study <- 
  ggplot()+
  geom_segment(aes(x = 0, y = 0, xend = 3, yend = 2))+
  geom_segment(aes(x = 3, y = 2, xend = 6, yend = 2))+
  annotate("text", x = 1.5, y = 0, label = "Accrual Time = 3") +
  annotate("text", x = 4.5, y = 0, label = "Minimum Follow-up Time = 3") +
  geom_segment(aes(x = 3, y = 0, xend = 3, yend = 2), linetype = "dashed")+
  annotate("text", x = 3, y = -0.1, label = "Study duration = 6") +
  geom_point(aes(x = 6, y = 2))+
  labs(x = "Time in years", y = "Number of patients") +
  scale_x_continuous(
      breaks = c(0, 3, 6),
      minor_breaks = 0:6,
    )  +
  scale_y_continuous(
    breaks = NULL,
    minor_breaks = NULL,
    expand = c(0, 0.1)
  )
```

To isolate the impact of the statistical method and software implementation, the following parameters were held constant across all scenarios:

::: {.columns}
::: {.column}

- **Accrual duration:** 3 years  
- **Additional follow-up after end of accrual:** 3 years   
- **Test sidedness:** two-sided logrank test  

Accrual was assumed to be uniform over the accrual period. Censoring arose solely from administrative study termination; no additional loss to follow-up, noncompliance, or competing risks were assumed.
:::
::: {.column}
::: {.light-content}
```{r}
#| out-width: 100%
p_study + theme_flatly(base_size = 15)
```
:::

::: {.dark-content}
```{r}
#| out-width: 100%
p_study + theme_darkly(base_size = 15)
```
:::
:::
:::
<!-- end columns -->

### Relevancy classification of design scenarios

Each scenario was assigned a relevancy category based solely on the Type I error rate and power, following the same rule‑based system used for two‑arm designs:

*   **High relevance**
    $$
      0.02 < \alpha < 0.15 \quad\text{and}\quad \text{power} > 0.75.
    $$
    These represent realistic single‑arm designs commonly used in phase II studies, especially in rare‑disease contexts.

*   **Medium relevance**  
    Scenarios not meeting the criteria for high or low relevance; these represent plausible but less standard design choices.

*   **Low relevance**
    $$
      \alpha > 0.20 
      \quad\text{or}\quad 
      \text{power} < 0.60 
      \quad\text{or}\quad 
      \text{power} \ge 0.99.
    $$
    These combinations are rarely used in practice but included intentionally to explore numerical robustness across asymptotic and exact methods.

The classification is descriptive only and does not filter or influence sample size computations.

### Scope of the comparison

For each of the 400 scenarios, all methods and software implementations were used to compute:

1. The **required number of events** to achieve the target power under the specified hazard ratio, and  
2. The corresponding **total sample size**, accounting for accrual and follow-up.

The design space was intentionally chosen to span both conventional trial settings and extreme configurations, allowing a comprehensive assessment of agreement, divergence, and stability across methods.

## Results 

Methods compared:

- nQuery : SOT1 / One Sample Log-Rank Test with Accrual
- `OneArm2stage::phase2.TTE()`^[Our rewrite that accepts accrual time (see [Fork-repo](https://github.com/thomas-arocena/OneArm2stage/tree/accrual-time) for more details)]
- `SampleSizeSingleArmSurvival::calcSampleSizeArcsine()` 
- `rashnu::oneSurvSampleSize()`

```{r}
x <- read_rds("../../../outputs/comp_surv_one_arm.rds")
```

```{r}
datatable(x$n_ratio)
```

::: {.light-content}
```{r}
#| out-width: 100%
gt_theme_flatly(x$tables$table_n_ratio)
x$plots$p_n_ratio + theme_flatly()
```
:::

::: {.dark-content}
```{r}
#| out-width: 100%
gt_theme_darkly(x$tables$table_n_ratio) 
x$plots$p_n_ratio + theme_darkly()
```
:::

- Across all relevancy levels, **none** of the evaluated methods reproduce nQuery’s sample sizes closely. Even our rewrite of `OneArm2Stage` (expected to match nQuery) shows systematic deviations, typically **below 1.00** in high‑ and medium‑relevancy settings.
- The `rashnu` and `sssas` methods implement the approach of @Nagashima_Noma_Sato_Gosho_2020 based on an arcsin transformation of the Kaplan–Meier estimator. These two methods consistently yield **larger** sample sizes than nQuery, resulting in N‑ratios frequently well **above 1.00**, sometimes substantially so.
- In **high‑relevancy** designs, all three methods already differ from nQuery, with `OneArm2Stage` underestimating and the Nagashima‑based methods overestimating. Dispersion is wide and no method stays close to the reference band.
- In **medium** and **low‑relevancy** scenarios, discrepancies become even larger. Extreme deviations are driven by designs where nQuery outputs very large sample sizes, amplifying differences between formulas and underlying assumptions.
- Overall pattern:  
  - **OneArm2Stage < nQuery** (systematic underestimation)  
  - **Nagashima‑based methods > nQuery** (systematic overestimation)  
  - **No method aligns tightly with nQuery**.

**Overall:** unlike the two‑arm survival settings, the one‑arm fixed survival framework shows **major method‑specific differences**, with no implementation providing close agreement with nQuery across realistic or extended design regions.

### Interactive results

You can explore all results interactively, filter designs, compare methods, and inspect individual cases in the interactive section of the website.

**➡️ [*“Interactive results”*](../../interactive/interactive.qmd)**