---
title: "Two-Arm Fixed Design with Binary endpoints"
bibliography: ../../references.bibtex
---

```{r}
#| label: setup
source("../results-setup.R")
```

## Design space explored

All comparisons were performed for a **two-arm, fixed-design randomized clinical trial** with a **binary primary endpoint**. The objective was to compare **sample size calculations** across commonly used software implementations over a broad but controlled design space, assuming a test for a **difference in proportions** under a pooled-variance framework.

### Varying design parameters

Four key design parameters were varied systematically:

-   **Type I error rate ($\alpha$, two-sided):**  
    $\alpha \in {0.01,\ 0.05,\ 0.10,\ 0.20,\ 0.49}$

-   **Power ($1 - \beta$):**  
    $1 - \beta \in {0.51,\ 0.80,\ 0.90,\ 0.99}$

-   **Control-group response rate:**  
    $\pi_c \in {0.10,\ 0.30,\ 0.50,\ 0.80,\ 0.90}$

-   **Treatment effect (difference in proportions):**  
    $\Delta\pi = \pi_e - \pi_c \in {0.05,\ 0.15,\ 0.25,\ 0.49}$

The experimental-group response rate was derived as  
$\pi_e = \pi_c + \Delta\pi$,  
with scenarios resulting in $\pi_e > 1$ discarded.

These parameters yield **300 valid design scenarios** in the full factorial combination.

### Fixed design assumptions

To isolate differences due purely to the statistical method or software implementation, the following characteristics of the design were held constant across scenarios:

-   **Allocation ratio (experimental : control):** 1 : 1
-   **Test sidedness:** two-sided

### Relevancy classification of design scenarios

To help interpret method agreement across the full parameter grid, each scenario was classified *a priori* according to its **practical relevance**.  
The classification follows a rule‑based system determined solely by the Type I error rate and power:

-   **High relevance**  
    Scenarios with
    $$
      0.02 < \alpha < 0.15 \quad\text{and}\quad \text{power} > 0.75,
    $$
    representing operating characteristics commonly used in realistic confirmatory settings.

-   **Medium relevance**  
    Scenarios not meeting the criteria for “high” or “low.”  
    These correspond to plausible but less conventional values of α and power used in exploratory or constrained designs.

-   **Low relevance**  
    Scenarios with
    $$
      \alpha > 0.20 
      \quad\text{or}\quad 
      \text{power} < 0.60 
      \quad\text{or}\quad 
      \text{power} \ge 0.99.
    $$
    These configurations are rarely used operationally (e.g., extremely liberal α or near‑perfect power), but are retained to evaluate numerical behavior in challenging cases.

The classification does **not** filter scenarios nor influence computation; it simply contextualizes discrepancies between methods across a wide design space.

### Scope of the comparison

For each of the **300 design scenarios**, all software tools and methods were used to compute the **required total sample size** to achieve the target power under the specified proportions.

The design space was intentionally chosen to span both conventional trial settings and extreme configurations, allowing a comprehensive assessment of agreement, divergence, and stability across methods.

## Z-pooled

### Methods compared

- East: Difference of Proportions test [PN-2S-DI]
- nQuery : PTT36 / Inequality Tests for Difference of Two Proportions
- `rpact::getSampleSizeRates()`

### Results

**Overall:** for two‑arm binary fixed‑design trials using the Z‑pooled test, `nQuery` and `rpact` reproduce East’s sample sizes with **near‑perfect consistency**, with meaningful differences only in edge‑case, low‑practicality scenarios.

- In **high‑ and medium‑relevancy** designs, both `nQuery` and `rpact` show **excellent agreement** with East, with N‑ratios consistently centered around 1.00 and virtually no dispersion.
- Even in the **low‑relevancy** region, the two methods remain very close to 1.00. A few isolated deviations appear, but they correspond to designs with extreme proportions or highly unbalanced variance assumptions.
- The plot confirms this stability: points are tightly clustered at 1.00 across the full range of East sample sizes, with only a small number of outliers in the low‑relevancy panel.
- No systematic bias is visible for either method, and neither method shows drift with increasing sample size.

```{r}
x_pooled <- read_rds("../../../outputs/comp_bin_fixed_pooled.rds")
```

::: {.light-content}
```{r}
#| out-width: 100%
gt_theme_flatly(x_pooled$tables$table_n_ratio)
x_pooled$plots$p_n_ratio + theme_flatly()
```
:::

::: {.dark-content}
```{r}
#| out-width: 100%
gt_theme_darkly(x_pooled$tables$table_n_ratio) 
x_pooled$plots$p_n_ratio + theme_darkly()
```
:::

## Exact 

### Methods compared

- East: Difference of Proportions test [PN-2S-DI]
- nQuery : PTT36 / Inequality Tests for Difference of Two Proportions
- `bbssr::BinarySampleSize(test = "Fisher")`

### Results

📌**Overall:** for exact tests in two‑arm binary designs, `bbssr` aligns well with nQuery, while East produces consistently smaller sample sizes.

- Using **nQuery as the reference**, `bbssr` stays reasonably close across all relevancy levels, with N‑ratios typically fluctuating around 1.00 and stabilising as sample sizes increase.
- East, however, shows **systematic underestimation** relative to nQuery.
- In **high** and **medium‑relevancy** scenarios, `bbssr` converges smoothly toward the reference, while East remains consistently below 1.00.
- In the **low‑relevancy** region, both methods show more dispersion, but the general pattern remains unchanged:  
    - `bbssr` ≈ nQuery  
    - **East < nQuery** (structural methodological difference)
- These differences arise from the underlying exact‑test implementations, see more at ["What Do East and nQuery Compute *Exactly* ?"](../../additional-content/bin2arms-exact.qmd).

```{r}
x_exact <- read_rds("../../../outputs/comp_bin_fixed_exact.rds")
```

::: {.light-content}
```{r}
#| out-width: 100%
gt_theme_flatly(x_exact$tables$table_n_ratio)
x_exact$plots$p_n_ratio + theme_flatly()
```
:::

::: {.dark-content}
```{r}
#| out-width: 100%
gt_theme_darkly(x_exact$tables$table_n_ratio) 
x_exact$plots$p_n_ratio + theme_darkly()
```
:::

## Interactive results

You can explore all results interactively, filter designs, compare methods, and inspect individual cases in the interactive section of the website.

**➡️ [*“Interactive results”*](../../interactive/interactive.qmd)**