---
title: "Two-Arm Group-Sequential-Design with Binary endpoints"
bibliography: ../../references.bibtex
---

```{r}
#| label: setup
source("../results-setup.R")
```

## Design space explored

All comparisons were performed for a **two‑arm, group‑sequential randomized clinical trial** with a **binary primary endpoint**. The aim was to evaluate the consistency of **sample size calculations** across commonly used software tools, over the same factorial design space used in the fixed‑design setting.

### Varying design parameters

The same four design parameters as in the fixed-design section were varied:

-   **Type I error rate ($\alpha$, two-sided):**  
    $\alpha \in {0.01,\ 0.05,\ 0.10,\ 0.20,\ 0.49}$

-   **Power ($1 - \beta$):**  
    $1 - \beta \in {0.51,\ 0.80,\ 0.90,\ 0.99}$

-   **Control-group response rate:**  
    $\pi_c \in {0.10,\ 0.30,\ 0.50,\ 0.80,\ 0.90}$

-   **Treatment effect (difference in proportions):**  
    $\Delta\pi \in {0.05,\ 0.15,\ 0.25,\ 0.49}$

The experimental-group rate was defined as  
$\pi_e = \pi_c + \Delta\pi$,  
and scenarios with $\pi_e > 1$ were excluded.

A total of **300 valid design scenarios** were obtained.

### Group-sequential design assumptions

To isolate differences attributable to software implementation and numerical methods, the group-sequential structure was fixed across all scenarios:

-   **Number of analyses:** 4 looks
-   **Spacing:** equally spaced by **information rate**
-   **Type I error spending:**  
    **Lan–DeMets O’Brien–Fleming–like** α‑spending function
-   **Power spending (futility):**  
    **Lan–DeMets O’Brien–Fleming–like** β‑spending, **non‑binding**
-   **Test:** two-sided test for a **difference in proportions (pooled Z-statistic)**
-   **Allocation ratio:** 1 : 1

Stopping for futility followed a **non‑binding** rule so that crossing the futility boundary does not inflate the Type I error.

### Relevancy classification of design scenarios

As in the fixed‑design setting, each scenario was assigned a relevancy category based solely on α and power:

*   **High relevance**
    $$
      0.02 < \alpha < 0.15 \quad\text{and}\quad \text{power} > 0.75.
    $$
    These settings correspond to realistic, routinely used confirmatory designs—now extended to the group-sequential framework.

*   **Medium relevance**  
    Scenarios falling outside the high and low regions, representing plausible but less typical combinations of α and power encountered in exploratory or early‑phase program designs.

*   **Low relevance**
    $$
      \alpha > 0.20 
      \quad\text{or}\quad 
      \text{power} < 0.60 
      \quad\text{or}\quad 
      \text{power} \ge 0.99.
    $$
    These extreme configurations are not typical for group‑sequential confirmatory trials but are kept intentionally to examine numerical robustness of spending‑function implementations.

The categorization is descriptive and plays no role in filtering or weighting the results; it serves purely to guide interpretation.

## Results

Method compared:

- East: Difference of Proportions test [PN-2S-DI]
- `rpact::getSampleSizeRates(design = rpact::getDesignGroupSequential)`


```{r}
x_pooled <- read_rds("../../../outputs/comp_bin_gs_pooled.rds")
```

::: {.light-content}
```{r}
#| out-width: 100%
gt_theme_flatly(x_pooled$tables$table_n_ratio)
x_pooled$plots$p_n_ratio + theme_flatly()
```
:::

::: {.dark-content}
```{r}
#| out-width: 100%
gt_theme_darkly(x_pooled$tables$table_n_ratio) 
x_pooled$plots$p_n_ratio + theme_darkly()
```
:::

- Across **all relevancy levels**, `rpact` shows **excellent agreement** with East, with N‑ratios clustering tightly around 1.00.
- In **high** and **medium‑relevancy** scenarios, variation is minimal: nearly all designs fall within a very narrow band (≈0.98–1.02), indicating practical equivalence.
- In the **low‑relevancy** region, a small number of mild deviations appear, but no systematic trend is observed. These differences stem from edge‑case parameter combinations rather than method behaviour.

**Overall:** `rpact` reproduces East’s group‑sequential Z‑pooled binary sample sizes with **near‑perfect consistency**, including in extended design regions.

### Interactive results

You can explore all results interactively, filter designs, compare methods, and inspect individual cases in the interactive section of the website.

**➡️ [*“Interactive results”*](../interactive/interactive.qmd)**