---
title: Comparison of Sample Size Calculations Across Methods
subtitle: R as an alternative to East and nQuery
format: clean-revealjs
author:
  - name: Arocena Thomas
    orcid: 0009-0002-7014-220X
    email: arocena.thomas@gmail.com
    affiliations: Oncostat
date: last-modified
toc-depth: 1
toc: true
---

```{r}
#| echo: false
#| message: false
#| warning: false
library(tidyverse)
```

# Introduction {background-color="#202B3E"}

## Motivations 
### Why R for sample size calculation ?

- do all analysis in R without having to use an external tool
- share code to someone (better than share an East of nQuery project)
- avoid problem with export
- free and open-source code, allows to open issue, ask new feature, fork the code

-> R packages need validation to get the same credibility and trust as proprietary softwares like East and nQuery

## Motivations 
### Why a comparision between R packages and East and nQuery ?

- East and nQuery already have great credibility and trust
- East and nQuery are widely used among statistian, especially here at Oncostat
- Prove that for a same design proprietary softwares and R packages give quite similar results derivates that trust into those packages.

## Scope of the comparison
### Explored clinical trials
```{mermaid}
flowchart TB
  A[Clinical Trial Design] --> AA[Time-to-Event Endpoints]
    AA --> AAA[Two-Arms]
      AAA --> AAAA[Fixed Design]
      AAA --> AAAB[Group-Sequential Design]
    AA --> AAB[One Arm]
  A --> AB[Binary Endpoints]
    AB --> ABA[Two-Arms]
      ABA --> ABAA[Fixed Design]
        ABAA --> ABAAA[Exact Computation]
        ABAA --> ABAAB[Z-pooled]
        ABAA --> ABAAC[Z-unpooled]
      ABA --> ABAB[Group-Sequential Design]
        ABAB --> ABABA[Z-pooled]
    AB --> ABB[One-Arm]
      ABB --> ABBA[Exact Computation]
      ABB --> ABBB[Non-exact Computation]
```

## Scope of the comparison {.smaller}
### Explored methods per clinical trials

| Endpoints  | Design Type              | Computation | Software           | R package                        |
|:-----------|:-------------------------|:------------|:-------------------|:---------------------------------|
| Survival   | Two-arm Fixed            |             | 🟧East, 🟦nQuery  | 🟥Rpact, 🟪Rashnu, 🟫gsDesign2 |
| Survival   | One-arm Fixed            |             | 🟦nQuery           | oa2s, sssas, 🟪Rashnu           |
| Survival   | Two-arm Group-Sequential |             | 🟧East, 🟦nQuery  | 🟥Rpact, 🟫gsDesign2            |
| Binary     | Two-arm Fixed            | Exact       | 🟧East, 🟦nQuery  | 🟨bbssr                         |
| Binary     | Two-arm Fixed            | Pooled      | 🟧East, 🟦nQuery  | 🟥Rpact, 🟨bbssr                |
| Binary     | Two-arm Fixed            | Unpooled    | 🟧East, 🟦nQuery  |                                  |
| Binary     | One-arm Fixed            | Exact       | 🟧East            | A'Hern(no package)               |
| Binary     | One-arm Fixed            | Non_exact?  | 🟧East, 🟦nQuery  | 🟥Rpact                         |
| Binary     | Two-arm Group-Sequential | Pooled      | 🟧East            | 🟥Rpact                          |

: Methods compared for each design

# Time-to-event endpoints {background-color="#1B7D7B"}

```{mermaid}
flowchart TB
AA[Time-to-Event Endpoints] --> AAA[Two-Arms]
  AAA --> AAAA[Fixed Design]
  AAA --> AAAB[Group-Sequential Design]
AA --> AAB[One Arm]
```

## [{{< fa circle >}}]{style="color: #202B3E; font-size: 0.5em;"} Quick theorical background
### Required input

- To compute required number of event we need:
  - $\alpha$: Type I error
  - power ($1-\beta$): jsp
  - hazard ratio: $\lambda_2/\lambda_1$

- In addition, to compute sample size we need:
  - Survival probability of control group at a certain time $t$
  - Accrual time
  - Follow-up time
  - Dropout rate

## Quick theorical background
### Schoenfeld formula

$$
e = \frac{4 (z_{1-\alpha/2} + z_{1-\beta})^2}{(\log(HR))^2}
$$

Other formula "Freedman" and "Hsieh-Freedman"


## Quick theorical background
### From required events to sample size

Basic idea : $n = e/\mathbb{P}(event)$

## Two-Arm fixed design
### Methods compared

| Endpoints  | Design Type              | Software           | R package                        |
|:-----------|:-------------------------|:-------------------|:---------------------------------|
| Survival   | Two-arm Fixed            | 🟧East, 🟦nQuery  | 🟥Rpact, 🟪Rashnu, 🟫gsDesign2 |

I would like to put the names of each functions

East : STT0 (no the good one)
nQuery : ARO3FJ

```{r}
#| echo: true
#| eval: false
rpact::getSampleSizeSurvival()
rashnu::lakatosSampleSize()
gsDesign2::fixed_design_ahr() # needs to define enroll_rate, fail_rate before
```

## Two-Arm fixed design
### Moving parameters

- **Type I error rate ($\alpha$, two-sided):**  
  $\alpha \in \{0.01, 0.05, 0.10, 0.20, 0.49\}$

- **Power ($1 - \beta$):**  
  $1 - \beta \in \{0.51, 0.80, 0.90, 0.99\}$

- **Hazard ratio (HR):**  
  $\text{HR} \in \{0.10, 0.50, 0.70, 0.90, 0.99\}$

- **Control-group survival probability at 3 years:**  
  $S_c(3\text{y}) \in \{0.10, 0.30, 0.60, 0.90\}$

## Two-Arm fixed design
### Additional parameters

:::: {.columns}

::: {.column width="50%"}

- **Accrual duration:** 3 years  
- **Additional follow-up after end of accrual:** 3 years   
- **Allocation ratio (experimental : control):** 1 : 1  
- **Test sidedness:** two-sided logrank test  

:::

::: {.column width="50%"}

```{r}
ggplot()+
geom_segment(aes(x = 0, y = 0, xend = 3, yend = 2))+
geom_segment(aes(x = 3, y = 2, xend = 6, yend = 2))+
annotate("text", x = 1.5, y = 0, label = "Accrual Time = 3") +
annotate("text", x = 4.5, y = 0, label = "Minimum Follow-up Time = 3") +
geom_segment(aes(x = 3, y = 0, xend = 3, yend = 2), linetype = "dashed")+
annotate("text", x = 3, y = -0.1, label = "Study duration = 6") +
geom_point(aes(x = 6, y = 2))+
labs(x = "Time in years", y = "Number of patients") +
scale_x_continuous(
    breaks = c(0, 3, 6),
    minor_breaks = 0:6,
  )  +
scale_y_continuous(
  breaks = NULL,
  minor_breaks = NULL,
  expand = c(0, 0.1)
) +
theme_classic()
```

:::

::::

## Two-Arm fixed design
### Results I

Some graphs

## Two-Arm fixed design
### Results II

Some other graphs

## Two-Arm Group sequential design
### Group-sequential design meaning

Interim analysis
We consider 4 looks equally spaced by information

- 25% of events
- 50% of events
- 75% of events
- 100% of events

## Two-Arm Group sequential design
### Lan & DeMets Spending function

What is a spending function, why ?
We choose alpha-spending function and 
beta spending function with O'Brien Flemming -like spending

## Two-Arm Group sequential design
### Compared method
| Endpoints  | Design Type              | Software           | R package                        |
|:-----------|:-------------------------|:-------------------|:---------------------------------|
| Survival   | Two-arm Group-Sequential | 🟧East, 🟦nQuery  | 🟥Rpact, 🟫gsDesign2            |

East : JEANBOB

nQuery : MICHMICHEL

```{r}
#| echo: true
#| eval: false
rpact::getDesignGroupSequential()
rpact::getSampleSizeSurvival() 
gsDesign2::gs_design_ahr()
```

## Two-Arm Group sequential design
### Results I

Some results 

## Two-Arm Group sequential design
### Results II

Some other results

## One-Arm fixed design
### Compared method
| Endpoints  | Design Type              | Software           | R package                        |
|:-----------|:-------------------------|:-------------------|:---------------------------------|
| Survival   | One-arm Fixed            | 🟦nQuery          | oa2s^[OneArm2stage], sssas^[SampleSizeSingleArmSurvival], 🟪Rashnu         |

nQuery : ROBERTJEAN

```{r}
#| echo: true
#| eval: false
OneArm2stage::phase2.TTE()
SampleSizeSingleArmSurvival::calcSampleSizeArcsine()
rashnu::oneSurvSampleSize()
```

## One-Arm fixed design
### Results I

Some results 

## One-Arm fixed design
### Results II

Some other results

<!-- # Binary endpoints {background-color="#82153F"}  -->
# **Binary endpoints**

[{{< fa circle >}}]{style="color: #82153F; font-size: 3em;"}


## [{{< fa circle >}}]{style="color: #82153F; font-size: 0.5em;"} Quick theorical background
### Required input

- To compute sample size we need:
  - $\alpha$: Type I error
  - power ($1-\beta$): jsp
  - Expected response proportion in control group
  - Expected response proportion in experimental group

## [{{< fa circle >}}]{style="color: #82153F; font-size: 0.5em;"} Quick theorical background
### Different computation for variance

Pooled, Unpooled, Exact computation
Binomial approximation ?


# Two-Arms Fixed Design

[{{< fa circle >}}]{style="color: #82153F; font-size: 3em;"}
[{{< fa circle >}}]{style="color: #E4B43A; font-size: 3em;"}

## [{{< fa circle >}}]{style="color: #82153F; font-size: 0.5em;"}[{{< fa circle >}}]{style="color: #E4B43A; font-size: 0.5em;"} Method compared 

| Endpoints  | Design Type              | Computation | Software           | R package                        |
|:-----------|:-------------------------|:------------|:-------------------|:---------------------------------|
| Binary     | Two-arm Fixed            | Exact       | 🟧East, 🟦nQuery  | 🟨bbssr                         |
| Binary     | Two-arm Fixed            | Pooled      | 🟧East, 🟦nQuery  | 🟥Rpact, 🟨bbssr                |
| Binary     | Two-arm Fixed            | Unpooled    | 🟧East, 🟦nQuery  |                                  |

EAST: efezfjizejf

NQUERY: '"gigjejrgk'

```{r}
#| echo: true
#| eval: false
bbssr::BinarySampleSize(Test = c('Chisq', 'Fisher', 'Fisher-midP', 'Z-pool', 'Boschloo'))
rpact::getSampleSizeRates()
```

## [{{< fa circle >}}]{style="color: #82153F; font-size: 0.5em;"}[{{< fa circle >}}]{style="color: #E4B43A; font-size: 0.5em;"} Two-Arms Fixed Design
### Results for Exact computation

Some results 

## [{{< fa circle >}}]{style="color: #82153F; font-size: 0.5em;"}[{{< fa circle >}}]{style="color: #E4B43A; font-size: 0.5em;"} Two-Arms Fixed Design
### Results pooled computation

Some other results 

## [{{< fa circle >}}]{style="color: #82153F; font-size: 0.5em;"}[{{< fa circle >}}]{style="color: #E4B43A; font-size: 0.5em;"} Two-Arms Group-Sequential Design
### Method compared 

| Endpoints  | Design Type              | Computation | Software           | R package                        |
|:-----------|:-------------------------|:------------|:-------------------|:---------------------------------|
| Binary     | Two-arm Group-Sequential | Pooled      | 🟧East            | 🟥Rpact, can add gsdesign2       |

EAST: efezfjizejf

NQUERY: '"gigjejrgk'

```{r}
#| echo: true
#| eval: false
rpact::getSampleSizeRates()
gsDesign2::gs_design_rd() # <- not done for the moment
```

## [{{< fa circle >}}]{style="color: #82153F; font-size: 0.5em;"}[{{< fa circle >}}]{style="color: #E4B43A; font-size: 0.5em;"} Two-Arms Group-Sequential Design
### Results

This is results

# One-Arm fixed Design
[{{< fa circle >}}]{style="color: #82153F; font-size: 3em;"}[{{< fa circle >}}]{style="color: #58A550; font-size: 3em;"}

## [{{< fa circle >}}]{style="color: #82153F; font-size: 0.5em;"}[{{< fa circle >}}]{style="color: #58A550; font-size: 0.5em;"} One-Arm fixed Design
### Method compared 

| Endpoints  | Design Type              | Computation | Software           | R package                        |
|:-----------|:-------------------------|:------------|:-------------------|:---------------------------------|
| Binary     | One-arm Fixed            | Exact       | 🟧East            | A'Hern(no package)               |
| Binary     | One-arm Fixed            | Non_exact?  | 🟧East, 🟦nQuery  | 🟥Rpact                         |

```{r}
#| echo: true
#| eval: false
rpact::getSampleSizeRates()
```

## One-Arm fixed Design
### Results

# Conclusion

## What we learnt

Not much

## Limitation

Each tools has it own limitations

## Two-Arm Fixed design
### Quick theorical background

Here we have some text that may run over several lines of the slide frame,
depending on how long it is.

- first item 
    - A sub item

Next, we'll brief review some theme-specific components.

- Note that _all_ of the standard Reveal.js
[features](https://quarto.org/docs/presentations/revealjs/)
can be used with this theme, even if we don't highlight them here.

## Additional theme classes

### Some extra things you can do with the clean theme

Special classes for emphasis

- `.alert` class for default emphasis, e.g. [important note]{.alert}.
- `.fg` class for custom colour, e.g. [important note]{.fg style="--col: #e64173"}.
- `.bg` class for custom background, e.g. [important note]{.bg style="--col: #e64173"}.

Cross-references

- `.button` class provides a Beamer-like button, e.g.
[[Summary]{.button}](#sec-summary)

## Want more?

### See our longer demo slides

We've deliberarely kept this template lean, to get you up and running as fast as
possible.

We provide a separate
[demo template](https://github.com/grantmcdermott/quarto-revealjs-clean-demo),
with more examples for integrating code, tables, figures, etc.

  - See the live demo slides
  [here](https://grantmcdermott.com/quarto-revealjs-clean-demo/template.html).

## Summary {#sec-summary}

### A minimalist and elegant presentation theme

The Quarto reveal.js clean theme aims to be a minimalist and elegant presention
theme. Here are some options to get you started:

Add the theme to an existing project.

```{.bash}
quarto install extension grantmcdermott/quarto-revealjs-clean
```

... or, create a new project using this slide deck as a lean template.

```{.bash}
quarto use template grantmcdermott/quarto-revealjs-clean
```

... or, create a new project using the demo slide deck as a full template.

```{.bash}
quarto use template grantmcdermott/quarto-revealjs-clean-demo
```
