---
title: "Two-Arm Group-sequential Design with Time-to-Event endpoints"
# bibliography: references.bibtex
---

```{r}
#| label: setup
source("../quarto-setup.R")
```

## Design space explored

Comparisons were conducted for a **two‑arm randomized clinical trial** with a **time‑to‑event primary endpoint**, evaluated using a **group‑sequential logrank test** under proportional hazards. The objective was to compare sample‑size and event‑count outputs across software implementations over a structured grid of design parameters.

### Varying design parameters

Four key parameters were varied systematically:

-   **Type I error rate ($\alpha$, two‑sided):** $\alpha \in \{0.01,\, 0.05,\, 0.10\}$
-   **Power ($1-\beta$):** $1-\beta \in \{0.70,\, 0.80,\, 0.90\}$
-   **Hazard ratio (HR):** $\text{HR} \in \{0.30,\, 0.60,\, 0.90\}$
-   **Control‑group survival probability at 3 years:** $S_c(3\text{y}) \in \{0.20,\, 0.50,\, 0.80\}$

The 3‑year survival probability was used to calibrate the baseline control‑group hazard rate assuming an exponential distribution, ensuring compatibility across methods requiring an explicit survival model.

The full factorial combination of these parameters yielded **81 distinct design scenarios**.

### Fixed design assumptions

```{r}
p_study <- 
  ggplot()+
  geom_segment(aes(x = 0, y = 0, xend = 3, yend = 2))+
  geom_segment(aes(x = 3, y = 2, xend = 6, yend = 2))+
  annotate("text", x = 1.5, y = 0, label = "Accrual Time = 3") +
  annotate("text", x = 4.5, y = 0, label = "Minimum Follow-up Time = 3") +
  geom_segment(aes(x = 3, y = 0, xend = 3, yend = 2), linetype = "dashed")+
  annotate("text", x = 3, y = -0.1, label = "Study duration = 6") +
  geom_point(aes(x = 6, y = 2))+
  labs(x = "Time in years", y = "Number of patients") +
  scale_x_continuous(
      breaks = c(0, 3, 6),
      minor_breaks = 0:6,
    )  +
  scale_y_continuous(
    breaks = NULL,
    minor_breaks = NULL,
    expand = c(0, 0.1)
  )
```

To isolate the impact of the statistical methodology and software implementation, the following parameters were held constant across all scenarios:

::: {.columns}
::: {.column}

-   **Number of interim looks:** $k = 4$
-   **Information fractions:** equally spaced
-   **Event time window:** 3 years
-   **Accrual period:** 3 years (uniform accrual)
-   **Post‑accrual follow‑up:** 3 years
-   **Allocation ratio (experimental : control):** 1 : 1
-   **Test sidedness:** two‑sided
-   **Censoring:** administrative only (no additional loss to follow‑up or competing risks)

:::
::: {.column}
::: {.light-content}
```{r}
#| out-width: 100%
p_study + theme_flatly(base_size = 15)
```
:::

::: {.dark-content}
```{r}
#| out-width: 100%
p_study + theme_darkly(base_size = 15)
```
:::
:::
:::
<!-- end columns -->

Group‑sequential boundaries were generated using **Lan–DeMets spending functions**:

-   **Alpha‑spending:** O’Brien–Fleming–like
-   **Beta‑spending:** O’Brien–Fleming–like, non‑binding.

### Relevancy classification of design scenarios

To aid interpretation of results across a wide parameter space, each design scenario was classified *a priori* according to its **practical relevance in confirmatory clinical trials**. Scenarios were categorized as **high**, **medium**, or **low relevance**, based on the extremeness of the design parameters.

- **High relevance** scenarios correspond to commonly used design choices in late-phase clinical trials (e.g. $\alpha \leq 0.10$, power 80–90%, HR in the range 0.5–0.7).
- **Medium relevance** scenarios represent extreme but plausible configurations that may arise in specific settings, such as rare diseases or early stopping considerations (e.g. $\alpha = 0.20$, HR $\leq 0.10$ or HR $\geq 0.90$).
- **Low relevance** scenarios include parameter combinations that are rarely used in practice (e.g. very high $\alpha$, near-null treatment effects, or power close to 50% or 100%) and were included primarily to assess the numerical behavior and robustness of asymptotic approximations.

This classification was not used to filter results but rather to **contextualize observed discrepancies** between methods and to distinguish clinically meaningful differences from purely theoretical or numerical artifacts.

### Scope of the comparison

For each of the 81 scenarios, all methods and software implementations were used to compute:

1. The **required number of events** to achieve the target power under the specified hazard ratio, and  
2. The corresponding **total sample size**, accounting for accrual and follow-up.

The design space was intentionally chosen to span both conventional trial settings and extreme configurations, allowing a comprehensive assessment of agreement, divergence, and stability across methods.

## Results

Methods compared: 

- East: Log Rank Test Given Accrual Duration and Study Duration [SU-2S-LRSD]
- `rpact::getSampleSizeSurvival()`
- `gsDesign2::gs_design_ahr()`

```{r}
x <- read_rds("../outputs/comp_surv_gs.rds")
```

::: {.light-content}
```{r}
#| out-width: 100%
gt_theme_flatly(x$tables$table_n_ratio)
x$plots$p_n_ratio + theme_flatly()
```
:::

::: {.dark-content}
```{r}
#| out-width: 100%
gt_theme_darkly(x$tables$table_n_ratio) 
x$plots$p_n_ratio + theme_darkly()
```
:::

- In **high‑relevancy** designs, both `gsdesign2` and `rpact` remain close to East, with N‑ratios typically between ~0.95 and 1.05. Small systematic shifts appear, but overall alignment is good for standard TTE GS setups.
- For **medium‑relevancy** scenarios, the dispersion increases and the curves reveal some parameter‑dependent patterns. Deviations remain moderate, but neither method stays perfectly centered around 1.00 across the full range.
- In **low‑relevancy** cases, discrepancies become more pronounced. Both methods deviate from the 1.00 baseline in opposite directions depending on the region, reflecting unstable behaviour in designs that are not operationally realistic.
- These differences are expected: Group-sequential Time-to-Event formulas are more sensitive to event‑timing assumptions, and the “low‑relevancy” scenarios include parameter combinations that would not be used in practice.
- `nQuery` is not included because the version used does **not** provide group‑sequential designs for Time-to-Event endpoints.

**Overall:** within realistic design regions, both `gsdesign2` and `rpact` reproduce East’s GS TTE sample sizes reasonably well, with wider divergence appearing only in non‑practical parameter regions.

### Interactive results

You can explore all results interactively, filter designs, compare methods, and inspect individual cases in the interactive section of the website.

**➡️ [*“Interactive results”*](../interactive.qmd)**
