---
title: "Two-Arm Fixed Design with Time-to-Event endpoints"
# bibliography: references.bibtex
---

```{r}
#| label: setup
source("../quarto-setup.R")
```

<!-- <details>
  <summary>Configuration requise</summary>
  <p>
    Nécessite un ordinateur équipé d'un système d'exploitation. L'ordinateur
    doit disposer d'une mémoire et, idéalement, d'un stockage à long terme. Un
    dispositif d'entrée ainsi qu'un dispositif de sortie sont recommandés.
  </p>
</details> -->

## Design space explored

All comparisons were conducted for a **two-arm, fixed-design randomized clinical trial** with a **time-to-event primary endpoint**, analyzed using the logrank test under a proportional hazards framework. The objective was to compare sample size and event calculations across commonly used software implementations over a broad but structured design space.

### Varying design parameters

Four key design parameters were varied systematically:

- **Type I error rate ($\alpha$, two-sided):**  
  $\alpha \in \{0.01, 0.05, 0.10, 0.20, 0.49\}$

- **Power ($1 - \beta$):**  
  $1 - \beta \in \{0.51, 0.80, 0.90, 0.99\}$

- **Hazard ratio (HR):**  
  $\text{HR} \in \{0.10, 0.50, 0.70, 0.90, 0.99\}$

- **Control-group survival probability at 3 years:**  
  $S_c(3\text{y}) \in \{0.10, 0.30, 0.60, 0.90\}$

The 3-year survival probability was used to calibrate the baseline hazard under an exponential survival assumption for the control group, ensuring consistency across methods requiring explicit specification of the underlying survival distribution.

The full factorial combination of these parameters yielded **400 distinct design scenarios**.

### Fixed design assumptions

To isolate the impact of the statistical method and software implementation, the following parameters were held constant across all scenarios:
```{r}
p_study <- 
  ggplot()+
  geom_segment(aes(x = 0, y = 0, xend = 3, yend = 2))+
  geom_segment(aes(x = 3, y = 2, xend = 6, yend = 2))+
  annotate("text", x = 1.5, y = 0, label = "Accrual Time = 3") +
  annotate("text", x = 4.5, y = 0, label = "Minimum Follow-up Time = 3") +
  geom_segment(aes(x = 3, y = 0, xend = 3, yend = 2), linetype = "dashed")+
  annotate("text", x = 3, y = -0.1, label = "Study duration = 6") +
  geom_point(aes(x = 6, y = 2))+
  labs(x = "Time in years", y = "Number of patients") +
  scale_x_continuous(
      breaks = c(0, 3, 6),
      minor_breaks = 0:6,
    )  +
  scale_y_continuous(
    breaks = NULL,
    minor_breaks = NULL,
    expand = c(0, 0.1)
  )
```

::: {.columns}
::: {.column}

- **Accrual duration:** 3 years  
- **Additional follow-up after end of accrual:** 3 years   
- **Allocation ratio (experimental : control):** 1 : 1  
- **Test sidedness:** two-sided logrank test  

Accrual was assumed to be uniform over the accrual period. Censoring arose solely from administrative study termination; no additional loss to follow-up, noncompliance, or competing risks were assumed.
:::
::: {.column}
::: {.light-content}
```{r}
#| out-width: 100%
p_study + theme_flatly(base_size = 15)
```
:::

::: {.dark-content}
```{r}
#| out-width: 100%
p_study + theme_darkly(base_size = 15)
```
:::
:::
:::
<!-- end columns -->

### Relevancy classification of design scenarios

To aid interpretation of results across a wide parameter space, each design scenario was classified *a priori* according to its **practical relevance in confirmatory clinical trials**. Scenarios were categorized as **high**, **medium**, or **low relevance**, based on the extremeness of the design parameters.

- **High relevance** scenarios correspond to commonly used design choices in late-phase clinical trials (e.g. $\alpha \leq 0.10$, power 80–90%, HR in the range 0.5–0.7).
- **Medium relevance** scenarios represent extreme but plausible configurations that may arise in specific settings, such as rare diseases or early stopping considerations (e.g. $\alpha = 0.20$, HR $\leq 0.10$ or HR $\geq 0.90$).
- **Low relevance** scenarios include parameter combinations that are rarely used in practice (e.g. very high $\alpha$, near-null treatment effects, or power close to 50% or 100%) and were included primarily to assess the numerical behavior and robustness of asymptotic approximations.

This classification was not used to filter results but rather to **contextualize observed discrepancies** between methods and to distinguish clinically meaningful differences from purely theoretical or numerical artifacts.

### Scope of the comparison

For each of the 400 scenarios, all methods and software implementations were used to compute:

1. The **required number of events** to achieve the target power under the specified hazard ratio, and  
2. Tthe corresponding **total sample size**, accounting for accrual and follow-up.

The design space was intentionally chosen to span both conventional trial settings and extreme configurations, allowing a comprehensive assessment of agreement, divergence, and stability across methods.

## Results 

Methods compared:

- East: Log Rank Test Given Accrual Duration and Study Duration [SU-2S-LRSD]
- nQuery : STT1 / Two Sample Log-Rank Test of Exponential Survival
- `rpact::getSampleSizeSurvival()`
- `rashnu::lakatosSampleSizeSurvival()`
- `gsDesign2::fixed_design_ahr()`


```{r}
x <- read_rds("../outputs/comp_surv_fixed.rds")
```

```{r}
# datatable(x$n_ratio)
```

::: {.light-content}
```{r}
#| out-width: 100%
gt_theme_flatly(x$tables$table_n_ratio)
x$plots$p_n_ratio + theme_flatly()
```
:::

::: {.dark-content}
```{r}
#| out-width: 100%
gt_theme_darkly(x$tables$table_n_ratio) 
x$plots$p_n_ratio + theme_darkly()
```
:::

- In **high‑relevancy** settings, all methods yield N‑ratios extremely close to 1.00, showing near‑perfect agreement with East for typical time‑to‑event designs.
- **Medium‑relevancy** scenarios introduce slightly more dispersion, but deviations remain small (generally within a few percent) and without systematic over‑ or under‑estimation.
- **Low‑relevancy** cases exhibit the largest discrepancies. These extreme designs make N‑ratios mechanically unstable and are not clinically meaningful.
- When ignoring these outlier regions, all four implementations again converge tightly around 1.00.
- Rashnu uses a Lakatos-based approach, but the difference in results is not significant here.(See [Lakatos and Lan](../lakatos_n_lan.qmd) post for more details)

**Overall:** within realistic design regions, the four methods provide **highly consistent fixed‑design TTE sample sizes** relative to East.

### Interactive results

You can explore all results interactively, filter designs, compare methods, and inspect individual cases in the interactive section of the website.

**➡️ [*“Interactive results”*](../interactive.qmd)**